---
title: "citibike_maps"
always_allow_html: true
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(shiny)
library(leaflet)
library(dplyr)
library(plotly)
library(ggplot2)
library(ggthemes)
library(gganimate)
library(maps)
library(dplyr)

library(rgdal)
library(sf)
library(magick)

library(viridis)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


```{r, read in manhattan rides data}
manhattan_rides_df <- read_csv("manhattan_rides.csv")

manhattan_rides_df <-
  manhattan_rides_df %>% 
  mutate(
    day_of_week = factor(day_of_week, ordered = T, 
                         levels = c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat")), 
    year = factor(year), 
    age_group = factor(age_group, ordered = T,
                       levels = c("18-25","26-35", "36-45", "46-55", "56-65", "66-85"))
  )

manhattan_rides_df %>% 
  group_by(age_group) %>% 
  summarize(min = min(age), max = max(age), obs = n())
```

```{r, read in hubs and hospital data}
hubs_df = 
  read_csv("hubs.csv") %>% 
  drop_na()

hospitals_df = 
  read_csv("hospitals.csv") %>% 
  drop_na()
```

```{r, side by side animation}
nyc = 
  as_tibble(
    map_data("state")
  ) %>% 
  filter(subregion == "manhattan")

bikes_2019 = manhattan_rides_df %>% 
  filter(year == "2019") %>% 
    group_by(stop_date, end_station_longitude, end_station_latitude) %>% 
  summarise(n_rides = n()) %>% 
  mutate(
    date = paste0("2019-", stop_date),
         date = as.Date(date, "%Y-%m-%d"))

bikes_2020 = manhattan_rides_df %>%
  filter(year == "2020") %>% 
    group_by(stop_date, end_station_longitude, end_station_latitude) %>% 
  summarise(n_rides = n()) %>% 
  mutate(
    date = paste0("2020-", stop_date),
         date = as.Date(date, "%Y-%m-%d"))

map_2019 = 
  ggplot() +
  geom_polygon(data = nyc, aes(x = long, y = lat, group = group), fill = "grey", alpha = 0.5) +
  geom_point(data = bikes_2019, aes(x = end_station_longitude, y = end_station_latitude, size = n_rides, color = n_rides), alpha = 0.5) +
  scale_size(range = c(2.5, 12.5), name = "Rides") +
  geom_point(data = hospitals_df, aes(x = Long, y = Lat), color = "red") + 
  geom_point(data = hubs_df, aes(x = Long, y = Lat), color = "green") +
  transition_time(date) +
  labs(title = "Date: {frame_time}", color = "Rides") +
  enter_grow() +
  exit_shrink() +
  ease_aes("sine-in-out")

animate(map_2019, duration = 25, fps = 5, width = 800, height = 800, renderer = gifski_renderer())
anim_save("map_2019.gif")

map_2020 = 
  ggplot() +
  geom_polygon(data = nyc, aes(x = long, y = lat, group = group), fill = "grey", alpha = 0.5) +
  geom_point(data = bikes_2020, aes(x = end_station_longitude, y = end_station_latitude, size = n_rides, color = n_rides), alpha = 0.5) +
  scale_size(range = c(2.5,12.5), name = "Rides") +
  geom_point(data = hospitals_df, aes(x = Long, y = Lat), color = "red") + 
  geom_point(data = hubs_df, aes(x = Long, y = Lat), color = "green") +
  transition_time(date) +
  labs(title = "Date: {frame_time}", color = "Rides") +
  enter_grow() +
  exit_shrink() +
  ease_aes("sine-in-out")

animate(map_2020, duration = 25, fps = 5, width = 800, height = 800, renderer = gifski_renderer())
anim_save("map_2020.gif")

#combining both animations using magick package
a_gif = animate(map_2019, width = 400, height = 400)
b_gif = animate(map_2020, width = 400, height = 400)

a_mgif = image_read(a_gif)
b_mgif = image_read(b_gif)

new_gif = image_append(c(a_mgif[1], b_mgif[1]))
for (i in 2:100) {
  combined = image_append(c(a_mgif[i], b_mgif[i]))
  new_gif = c(new_gif, combined)
}

new_gif

```

* gif still needs adjusting so it is not so glitchy-looking (need to slow down animation I think)
* also seems like the projection for the Manhattan polygon and the citibike data do not quite match up
* need to fix legend to show red dots = hospitals and green dots = hubs
* maybe try to find more interesting way to display hospitals and hubs?

```{r, rides in 2019 vs 2020 by month}
#In retrospect I probably should have written a function to do this
nyc = 
  as_tibble(
    map_data("state")
  ) %>% 
  filter(subregion == "manhattan")

bikes_month_2019 = manhattan_rides_df %>% 
  filter(year == "2019") %>% 
  mutate(
    month = format(starttime, format = "%m"),
    month = as.numeric(month)
  ) %>% 
  group_by(month, end_station_longitude, end_station_latitude) %>% 
  summarise(n_rides = n())

bikes_month_2020 = manhattan_rides_df %>%
  filter(year == "2020") %>% 
  mutate(
    month = format(starttime, format = "%m"),
    month = as.numeric(month)
  ) %>%
  group_by(month, end_station_longitude, end_station_latitude) %>% 
  summarise(n_rides = n())

map_month_2019 = 
  ggplot() +
  geom_polygon(data = nyc, aes(x = long, y = lat, group = group), fill = "grey", alpha = 0.5) +
  geom_point(data = bikes_month_2019, aes(x = end_station_longitude, y = end_station_latitude, size = n_rides, color = n_rides), alpha = 0.5) +
  scale_size(range = c(1,8), name = "Rides") +
  geom_point(data = hospitals_df, aes(x = Long, y = Lat), color = "red") + 
  geom_point(data = hubs_df, aes(x = Long, y = Lat), color = "green") +
  transition_time(month) +
  labs(title = "2019 Month: {frame_time}", color = "Rides") +
  enter_grow() +
  exit_shrink() +
  ease_aes("sine-in-out")

map_month_2020 = 
  ggplot() +
  geom_polygon(data = nyc, aes(x = long, y = lat, group = group), fill = "grey", alpha = 0.5) +
  geom_point(data = bikes_month_2020, aes(x = end_station_longitude, y = end_station_latitude, size = n_rides, color = n_rides), alpha = 0.5) +
  scale_size(range = c(1,8), name = "Rides") +
  geom_point(data = hospitals_df, aes(x = Long, y = Lat), color = "red") + 
  geom_point(data = hubs_df, aes(x = Long, y = Lat), color = "green") +
  transition_time(month) +
  labs(title = "2020 Month: {frame_time}", color = "Rides") +
  enter_grow() +
  exit_shrink() +
  ease_aes("sine-in-out")

#combining both animations using magick package
a_month_gif = animate(map_month_2019, duration = 12, fps = 1, width = 400, height = 400)
b_month_gif = animate(map_month_2020, duration = 12, fps = 1, width = 400, height = 400)

a_month_mgif = image_read(a_month_gif)
b_month_mgif = image_read(b_month_gif)

new_month_gif = image_append(c(a_month_mgif[1], b_month_mgif[1]))
for (i in 2:12) {
  combined_month = image_append(c(a_month_mgif[i], b_month_mgif[i]))
  new_month_gif = c(new_month_gif, combined_month)
}

new_month_gif
```

It works!!

```{r, exploring other map options}
leaflet() %>% 
   addTiles() %>% 
   setView( lng = -73.98776, lat = 40.74488, zoom = 10.5 ) %>% 
   addProviderTiles("CartoDB.Positron")
```

I think leaflet and plotly can work together, but then animation is getting into using shiny. Might be worth it?

GGPlot - Updated for bikes 2019
```{r, 2019 animation by month}
bike_map = manhattan_rides_df %>% 
  filter(year == "2019") %>% 
  mutate(
    month = format(starttime, format = "%m"),
    month = as.numeric(month)
  ) %>%
    group_by(month, end_station_longitude, end_station_latitude) %>% 
  summarise(n_rides = n()) %>% 
  ggplot() +
  geom_point(aes(x = end_station_longitude, y = end_station_latitude, size = n_rides), alpha = 0.5) + 
  transition_time(month) +
  labs(title = "Month: {frame_time}", color = "Station") +
  ease_aes("linear")

animate(bike_map, duration = 12, fps = 1, width = 800, height = 800, renderer = gifski_renderer())
anim_save("bike_map.gif")
```

So we can see that the number of rides increases in the summer months and decreases in winter months.

